---
title: "template"
subtitle: "Project title"
author: "Jacob J. Christensen"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
    code_folding: hide
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	eval = TRUE, 
  echo = TRUE, 
  message = TRUE, 
  warning = FALSE, 
	root.dir = normalizePath(".."),
	dev = c("png", "cairo_pdf"), 
	dpi = 300, 
  fig.align = "center", 
	fig.keep = "all", 
	fig.path = "../results/figures/...-"
)
```

## Load libraries, data and project functions

First, we load necessary *libraries*, *data* and general-use project *functions*. Also, we set the ggplot2 *theme* to a type that is a bit more publication-ready. 

```{r}
# Attach libraries and functions
pkgs <- c("tidyverse", "tidylog")
inst <- lapply(pkgs, library, character.only = TRUE)
source("../scripts/functions.R")

# Load data
# ...

# Set project theme
theme_set(theme_minimal() + theme(axis.line = element_line()))
```

## Info

> This is a template file. There are many like it, but this one is mine. 



## Topic 1

```{r}
```


```{r}
```


```{r}
```



## Topic 2

```{r}
```


```{r}
```


```{r}
```



## Topic 3

```{r}
```


```{r}
```


```{r}
```



## Conclusions

This concludes the script. Some conclusions and take-homes: 

- 
- 
- 



## Save

Here we save those files that we will use in downstream work. 

```{r}
# saveRDS(..., file = "../data/processed/INSERT-NAME.rds")
# saveRDS(..., file = "../data/processed/INSERT-NAME.rds")
# saveRDS(..., file = "../data/processed/INSERT-NAME.rds")
# saveRDS(..., file = "../data/processed/INSERT-NAME.rds")
# saveRDS(..., file = "../data/processed/INSERT-NAME.rds")
```



## Session info

To improve reproducibility, print out the session info for this script. 

```{r}
devtools::session_info()
```

