---
title: "Multivariate analyses: PCA for XXX....."
subtitle: "Project XXX....."
author: "Jacob J. Christensen"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
    code_folding: show
  pdf_document:
    toc: yes
    toc_depth: '5'
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	eval = TRUE, 
  echo = TRUE, 
  message = FALSE, 
  warning = FALSE, 
	root.dir = normalizePath(".."), 
	dev = c("png", "pdf"),
	fig.align = "center", 
	fig.keep = "all", 
	fig.path = "../figures/pca-XXX.....-"
)
```

## Load libraries, data and project functions

```{r}
# Attach libraries
pkgs <- c("tidyverse", "RColorBrewer", "broom", "factoextra", "dendextend")
inst <- lapply(pkgs, library, character.only = TRUE)

# Add functions
source("../scripts/functions.R")

# Load data
# data_input <- load(file = "../data/processed/data-input.rds")
# variables <- load(file = "../data/processed/data-input-vars.rds")

# Set project theme
theme_set(theme_minimal() + theme(axis.line = element_line()))
```

## Info

Data are multivariate in nature and needs to be seen as such. Therefore, I here go through *PCA and correlation-based clustering* for all the different datasets. I will try to understand the underlying structure of the data. 

In this file, I use the **XXX.....** dataset, and associated variables vector. 

> This entire file should run fine (all the way *from loading to saving*) by changing these few lines only: the *modify* chunk, the title, and the figure path in the knitr chunk.

```{r modify}
dataset_name <- "XXX....."
dataset <- data_XXX.....
variables <- variables_XXX.....
```


## PCA

### Run analysis

First, run PCA and extract results for samples, variables and eigenvalues in a tidy format.

```{r}
# Run PCA
dataset_pca <- dataset %>% 
  select(IDcomb, one_of(variables)) %>% 
  as.data.frame() %>% 
  column_to_rownames(var = "IDcomb") %>% 
  prcomp(~ ., data = ., scale = TRUE, center = TRUE)

# Extract results in tidy format
pca_samples <- tidy(dataset_pca, matrix = "samples")
pca_variables <- tidy(dataset_pca, matrix = "variables")
pca_eigenvalues <- tidy(dataset_pca, matrix = "pcs")
```


### Eigenvalues

First examine the eigenvalues to figure out how many PCs we need to explain 50-99% of the variation in the data.

```{r}
value <- c(0.5, 0.75, 0.90, 0.95, 0.99)
value_comp <- map_dbl(value, ~number_of_pcs(pca_eigenvalues, .x))
```

To explain 50, 75, 90, 95 and 99 % of the variation, you would need `r value_comp[1]`, `r value_comp[2]`, `r value_comp[3]`, `r value_comp[4]`, `r value_comp[5]` principal components, respectively. 

Next, print the standard deviation, and percent and cumulative percent for each PC. Highlight the selected thresholds, as defined above.

```{r scree, fig.width=8, fig.height=3}
pca_eigenvalues_selected <- pca_eigenvalues %>% 
  slice(value_comp) %>% 
  gather(-PC, key = "key", value = "value")
  
pca_eigenvalues %>% 
  gather(-PC, key = "key", value = "value") %>% 
  ggplot(aes(PC, value)) + 
  geom_line() + 
  geom_label(data = pca_eigenvalues_selected, aes(label = PC), 
             
             # make the label small, remove border, and remove exessive padding
             size = 2, label.size = 0, label.padding = unit(0.15, "lines")) + 
  facet_wrap(~ fct_rev(key), scales = "free") + 
  labs(x = "Number of principal components", y = NULL, 
       title = paste0("Number of PC needed to explain ", 
                      paste0(value*100, collapse = ", "), 
                      " % \nof the cumulative percentage explained variance"))
```



### PC1 vs PC2

Now, visualize PC1 versus PC2. 

```{r}
# Prep samples
pca_samples_selected <- pca_samples %>% 
  spread(PC, value) %>% 
  gather(-row, -`1`, key = "key", value = "value") %>% 
  filter(key %in% 2:5) %>% 
  mutate(key = paste0("PC", key)) %>% 
  rename(PC1 = `1`)

# Prep vars
pca_variables_selected <- pca_variables %>% 
  spread(PC, value) %>% 
  gather(-column, -`1`, key = "key", value = "value") %>% 
  filter(key %in% 2:5) %>% 
  mutate(key = paste0("PC", key)) %>% 
  rename(PC1 = `1`)
```


#### Samples

Make a simple sample PCA plot. You can color by a group or other feature and add ellipses using `stat_ellipse` and `directlabels::geom_dl` for example. 

```{r pca-samples, fig.width=5, fig.height=4}
plot_pca_samples <- pca_samples_selected %>% 
  filter(key == "PC2") %>% 
  spread(key = "key", value = "value") %>% 
  mutate(id = as.character(row) %>% as.numeric()) %>% 
  ggplot(aes(PC1, PC2)) + 
  geom_point() + 
  # stat_ellipse(type = "t", level = 0.8, linetype = 2) +
  # directlabels::geom_dl(aes(label = group), method = "smart.grid") +
  scale_color_brewer(palette = "Dark2") + 
  theme(legend.position = "none") +
  labs(x = paste0("PC1 (", round(pca_eigenvalues[1, 3]*100, 1), " %)"),
       y = paste0("PC2 (", round(pca_eigenvalues[2, 3]*100, 1), " %)"))
plot_pca_samples

# To combine multiple PCA plots (for different data types) into a single figure, save it to a RDS file and retrieve it in a separate .Rmd script
# saveRDS(plot_pca_samples, file = "../data/processed/pca-samples-XXX.rds")
```


#### Biplot

Now make a biplot directly, with variables underneath, and the samples overlayed.

```{r biplot, fig.width=12, fig.height=10}
pca_samples_selected %>% 
  ggplot(aes(scale(PC1), scale(value))) + 
  geom_text(data = pca_variables_selected, 
            aes(label = column, size = abs(PC1) + abs(value)), 
            color = "blue", alpha = 0.3) + 
  geom_point(shape = 21, color = "white", fill = "grey50", size = 3, alpha = 0.15) + 
  facet_wrap(~ key, scales = "free") + 
  theme(legend.position = "none") + 
  labs(x = "PC1", y = NULL, 
       title = "Biplot: PC1 versus PC2, PC3, PC4 and PC5")
```

There seems to be some clustering of variables across the different dimensions. Let's cluster the variables and color them.


## Correlation heatmap

Run correlation and visualize the correlation matrix: 

```{r pheatmap, fig.width=9, fig.height=7}
dataset %>% 
  select(one_of(variables)) %>% 
  cor(use = "pairwise.complete.obs", method = "spearman") %>% 
  pheatmap::pheatmap()
```

We see some correlation between variables, suggesting redundant information in the dataset.


## Correlation cluster analysis

Run correlation and then cluster the distance object:

```{r}
dataset_hclust <- dataset %>% 
  select(one_of(variables)) %>% 
  cor(use = "pairwise.complete.obs", method = "spearman") %>% 
  dist(method = "euclidean") %>% 
  hclust(method = "complete")
```

Get some nice colors for the plot:

```{r}
# display.brewer.all(n = NULL,
#                    type = "qual",
#                    select = NULL,
#                    exact.n = TRUE,
#                    colorblindFriendly = FALSE)
dark2 <- brewer.pal(n = 8, name = "Dark2")
set1 <- brewer.pal(n = 9, name = "Set1")
mycols <- c(dark2, set1)
```


### Standard viz

Visualize the tree, and color by group:

```{r dendro-initial, fig.width=15, fig.height=4}
dend <-  as.dendrogram(dataset_hclust) %>%
  # Branches line width
  set("branches_lwd", 1) %>% 
  # Color branches by groups
  set("branches_k_color", mycols, k = 10) %>% 
  # Color labels by groups
  set("labels_colors", mycols, k = 10) %>%  
  # Change label size
  set("labels_cex", 0.6) %>% 
  # Hang leaves
  set("hang_leaves")
fviz_dend(dend)
```


### Circular viz

Visualize the circular plot instead:

```{r dendro-circular, fig.width=10, fig.height=7}
fviz_dend(dend, type = "circular")
```


### Cluster biplot

Cut into 10 groups, to be able to visualize the variables in the PCA plot.

```{r}
hclust_clusters <- dataset_hclust %>% 
  stats::cutree(k = 10) %>% 
  tibble(variable = names(.), 
         cluster = factor(.)) %>% 
  select(-.)
```

Visualize the clusters again at the biplot.

```{r biplot-clust, fig.width=12, fig.height=10}
pca_variables_selected_clusters <- pca_variables_selected %>% 
  left_join(hclust_clusters, by = c("column" = "variable"))

pca_samples_selected %>% 
  ggplot(aes(scale(PC1), scale(value))) + 
  geom_point(shape = 21, color = "white", fill = "grey50", size = 3, alpha = 0.15) + 
  geom_text(data = pca_variables_selected_clusters, 
            aes(label = column, color = cluster, 
                size = abs(PC1) + abs(value), 
                alpha = abs(PC1) + abs(value))) + 
  facet_wrap(~ key, scales = "free") + 
  theme(legend.position = "none") + 
  labs(x = "PC1", y = NULL, 
       title = "Biplot: PC1 versus PC2, PC3, PC4 and PC5")
```




## Brief summary

...


## Save

Since most of this document is generic, I need to rename my objects prior to saving. To this, use the `assign` function.

Assign:

```{r}
# res <- list(
#   assign(paste0("pca_samples", "_", dataset_name), pca_samples), 
#   assign(paste0("pca_variables", "_", dataset_name), pca_variables), 
#   assign(paste0("pca_eigenvalues", "_", dataset_name), pca_eigenvalues), 
#   assign(paste0("hclust_clusters", "_", dataset_name), hclust_clusters)) %>% 
#   set_names(c(
#     paste0("pca_samples", "_", dataset_name), 
#     paste0("pca_variables", "_", dataset_name), 
#     paste0("pca_eigenvalues", "_", dataset_name), 
#     paste0("hclust_clusters", "_", dataset_name)))
```

```{r}
res <- list(pca_samples, pca_variables, pca_eigenvalues, hclust_clusters) %>% 
  set_names(c(
    paste0("pca_samples", "_", dataset_name),
    paste0("pca_variables", "_", dataset_name),
    paste0("pca_eigenvalues", "_", dataset_name),
    paste0("hclust_clusters", "_", dataset_name)))
```


Save:

```{r, include=FALSE}
# Effectively save PCA/cluster results to .rds objects in the ../data/processed/ directory; for use in the "pca-..." reports
save_with_name = function(name, object){
  
  # "name" is a string with the name of the variable you want to assign
  # "object" is any kind of R object (data.frame, list, etc.) you want to rename and save
  
  assign(name, object)
  save(list = name, file = sprintf("../data/processed/%s.rds", name))
}

save_with_name(name = sprintf("pca_%s", dataset_name), object = res)
```

## Session info

```{r session info, echo=FALSE}
devtools::session_info()
```
