---
title: "project name"
subtitle: "A substudy from the <project name>"
author: "Jacob J. Christensen"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
pkgs <- c("tidyverse", "rmarkdown", "knitr")
inst <- lapply(pkgs, library, character.only = TRUE)
opts_chunk$set(echo = TRUE)
```

This document is simply to run all project-related markdown documents effectively, in correct order, and to avoid having to repeatedly copy-paste figures and tables over to the documentation folder. 



## Paths

```{r}
manu <- "christensen-<project-name>"
path_res <- "../results"
path_res_figs <- "../results/figures"
path_res_tabs <- "../results/tables"
path_manu <- file.path("../doc", manu)
path_manu_figs <- file.path("../doc", manu, "figures")
path_manu_tabs <- file.path("../doc", manu, "tables")
path_manu_sub <- file.path("../doc", manu, "submissions")
```



## Script names

```{r}
list.files() %>% str_subset(".Rmd")
```



## Reports

```{r}
render_report <- function(input) {
  render(input = input, 
         output_format = "html_document", 
         output_dir = "../results/reports/")
  }
```
  
```{r}
render_report("01-XXX.Rmd")
render_report("02-XXX.Rmd")
render_report("03-XXX.Rmd")
render_report("04-XXX.Rmd")
```



## Manuscripts

### Main manuscript

```{r}
render(
  input = file.path(path_manu, "christensen-<project-name>-main.Rmd"), 
  output_format = "rmarkdown::word_document", 
  output_dir = path_manu)
```

### Supplemental material manuscript

```{r}
render(
  input = file.path(path_manu, "christensen-<project-name>-suppl.Rmd"), 
  output_format = "rmarkdown::word_document", 
  output_dir = path_manu)
```



## Figures

Which figures to submit? This is determined from the manuscript.

These are all the figure names: 

```{r}
list.files(path_res_figs)
```

### Figure names

```{r}
name_figs <- c(
  
    "figure-1" = "", # brief_name
    "figure-2" = "", # brief_name
    "figure-3" = "", # brief_name
    "figure-4" = "", # brief_name
    
    "figure-s1" = "", # brief_name
    "figure-s2" = "", # brief_name
    "figure-s3" = "", # brief_name
    "figure-s4" = "", # brief_name
    "figure-s5" = "", # brief_name
    "figure-s6" = "", # brief_name
    "figure-s7" = "" # brief_name
    
    )
```


### Copy to doc and rename

This is a function to move one figure from the results/figures to the doc/manuscript/figures/ folder, and rename it according to manuscript, as defined in the previous chunk.

Note that this implementation *overwrites* figures already in the doc/manu/figures folder. Therefore, at this point, it's a good idea to go over and see if there are any that shouldn't be present in the manuscript/figures folder.

```{r}
copy_and_rename_figures <- function(
  
  name_figs, 
  name_figs_new,
  overwrite = FALSE
  
  ) {
  
  if (length(list.files(path_manu_figs)) > 0 && isFALSE(overwrite)) 
    stop(paste0("Be careful not to overwrite existing files in the ", path_manu_figs, " directory!\n"))
  if (length(list.files(path_manu_figs)) > 0 && isTRUE(overwrite)) 
    warning(paste0("Be careful when you overwrite existing files in the ", path_manu_figs, " directory! 
                   \nNext time, go delete the figures that you don't need properly beforehand.\n\n"))

  # Copy figures
  name_figs %>% 
    map(~list.files(path_res, .x, full.names = TRUE) %>% str_subset(".png")) %>% 
    map(~file.copy(.x, path_manu_figs))
    
  name_figs %>% 
    map(~list.files(path_res_figs, .x, full.names = TRUE) %>% str_subset(".png|.pdf")) %>% 
    map(~file.copy(.x, path_manu_figs))
    
  # Get figure suffixes
  manu_figures <- name_figs %>% 
    map(~list.files(path_manu_figs, .x, full.names = TRUE))
  
  # Get figure extensions
  figure_extensions <- manu_figures %>% 
    map(~str_sub(.x, start = -4L, end = -1L))
  
  # Rename figures
  pmap(
    list(manu_figures, name_figs_new, figure_extensions), 
    ~file.rename(from = ..1, to = file.path(path_manu_figs, paste0(..2, ..3))))
}
```

Evaluate if you should delete figures already there; then, apply the function to all figures defined above.

```{r}
# R figures, flow charts and DAGs
copy_and_rename_figures(name_figs, names(name_figs), overwrite = FALSE)
```




## Tables

These are the main and supplemental table names:

```{r}
list.files(path = "../results/tables/")
```

### Table names

```{r}
names_tab <- c(
  "table-1" = "", # brief_name
  
  "table-s1" = "", # brief_name
  "table-s2" = "", # brief_name
  "table-s3" = "", # brief_name
  "table-s4" = "", # brief_name
  "table-s5" = "", # brief_name
  "table-s6" = "", # brief_name
  "table-s7" = "" # brief_name
)
```

### Gather and rename

Gather all tables to a single excel file.

```{r}
# Take the table names defined in the previous chunk
names_tab %>% 
  
  # Subset the full path of those names
  map(~str_subset(string = list.files(path = "../results/tables/", full.names = TRUE), pattern = .x)) %>% 
  
  # Read the tables
  map(read_csv) %>%
  
  # Add the names
  set_names(names(names_tab)) %>%
  
  # And write everything to a single excel file
  openxlsx::write.xlsx(file.path(path_res_tabs, "tables-all.xlsx"))
```

### Copy to doc

Evaluate if you should delete the excel tables already there; then, copy the updated version to doc/manuscript/tables folder.

```{r}
file.copy(file.path(path_res_tabs, "tables-all.xlsx"), 
          file.path(path_manu_tabs, "tables-all.xlsx"), overwrite = FALSE)
```




## Submissions

For each submission, we will do the following: Take the main and supplemental manuscript documents, figure and tables, etc, and copy into a type and date-specific submission folder.

List files and folders in the doc/manu directory:

```{r}
list.files(path_manu)
```

The "prepare submission" function:

```{r}
prep_submission <- function(type = NULL, journal = NULL, overwrite = FALSE) {
  
  # Test that type and journal is correct
  type <- match.arg(arg = type, choices = c("co-authors", "journals"))
  if (!is.null(journal)) stopifnot(is.character(journal))
  
  # Create path
  if (type == "co-authors") path_sub <- file.path(path_manu_sub, type, Sys.Date())
  else if (type == "journals") path_sub <- file.path(path_manu_sub, type, paste0(Sys.Date(), "-", journal))
  
  # Create dir at path
  dir.create(path_sub)
  
  # Take all files
  file.copy(
    c(list.files(path_manu, full.names = TRUE) %>% str_subset("-main.docx"), 
      list.files(path_manu, full.names = TRUE) %>% str_subset("-suppl.docx"),
      list.files(path_manu_tabs, full.names = TRUE), 
      list.files(path_manu_figs, full.names = TRUE)), 
    path_sub)
  
  # Zip everything for simple distribution
  zip::zipr(
    zipfile = paste0(file.path(path_sub, manu), ".zip"),
    files = list.files(path_sub, full.names = TRUE)
  )
}
```

Prepare submission: 

```{r}
prep_submission("co-authors")
# prep_submission("journal", "ajcn")
```




## Session info

```{r}
devtools::session_info()
```





---